---
title: "Qusage: Speeding up in Rcpp"
author:  "Timothy J. Triche, Jr, Anthony R. Colombo"
output: 
    pdf_document:
      toc: true
      number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

#SpeedSage Intro
qusage is published software that is slow for large runs, SpeedSage corrects for speed and efficiency at large orders
#Bottlenecking of Functions
Qusage can improve the speed of its algorithm by minimizing the cost of computaiton.

##changes calcIndividualExpressionsC
trading NA flexibility slows down qusage runs, but having the user input no NAs enforcing good input, this speeds up calcIndividualExpressionsC 2X

#Individual Expression Function
This test the local version which enforces no NA in Baseline or PostTreatment object, this reduces the flexibility. this test data is from the vignette where postTreatment was modified to be Baseline+20.4, a simple training set.

```{r} 
library(inline)
library(microbenchmark)
library(Rcpp)
library(parallel)
library(speedSage)
library(qusage)
library(ggplot2)
eset<-system.file("extdata","eset.RData",package="speedSage")
load(eset)
labels<-c(rep("t0",134),rep("t1",134))
contrast<-"t1-t0"
colnames(eset)<-c(rep("t0",134),rep("t1",134))
fileISG<-system.file("extdata","c2.cgp.v5.1.symbols.gmt",package="speedSage")
ISG.geneSet<-read.gmt(fileISG)
ISG.geneSet<-ISG.geneSet[grepl("DER_IFN_GAMMA_RESPONSE_UP",names(ISG.geneSet))]
Baseline<-eset
PostTreatment<-eset+20.4
#paired
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmaArm.cpp")
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmasCpp.cpp")
test1<-calcIndividualExpressionsArm(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6)
test2<-calcIndividualExpressionsC(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6)
test3<-calcIndividualExpressions(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6,na.rm=TRUE)
qplot(abs(test1[[1]]-test3[[1]]))
qplot(abs(test1[[2]]-test3[[2]]))
qplot(abs(test1[[3]]-test3[[3]]))
qplot(abs(test1[[4]]-test3[[4]]))

mb<-microbenchmark(
test1<-calcIndividualExpressionsArm(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6),
test2<-calcIndividualExpressionsC(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6),
test3<-calcIndividualExpressions(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6,na.rm=TRUE))
mb

require(profr)
require(ggplot2)
x1<-profr(calcIndividualExpressions(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6,na.rm=TRUE))
ggplot(x1)+labs(title="Qusage SE Default")
x2<-profr(calcIndividualExpressionsArm(Baseline,PostTreatment,paired=TRUE,min.variance.factor=10^-6))
ggplot(x2)+labs(title="Qusage SE Armadillo")



#single end testing
sourceCpp("/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmaSingle.cpp")
testSE1<-calcIndividualExpressions(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6,na.rm=TRUE)
testSE2<-calcIndividualExpressionsArm(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6)
testSE3<-calcIndividualExpressionsC(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6)


e1<-(abs(testSE1[[1]]-testSE2[[1]]))
e2<-(abs(testSE1[[2]]-testSE2[[2]]))
e3<-(abs(testSE1[[3]]-testSE2[[3]]))
e4<-(abs(testSE1[[4]]-testSE2[[4]]))
qplot(as.vector(e1))
qplot(as.vector(e2))
qplot(as.vector(e3))
qplot(as.vector(e4))



require(profr)
require(ggplot2)
y1<-profr(calcIndividualExpressions(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6,na.rm=TRUE))
y2<-profr(calcIndividualExpressionsArm(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6))
ggplot(y1)+labs(title="Qusage SE Default")
ggplot(y2)+labs(title="Qusage SE Arm")

#this shows that the only difference is the vector of Non-NA columns per each row; which is the same as the number of columns if no-na is enforced.
seMB<-microbenchmark(
testSE1<-calcIndividualExpressions(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6,na.rm=TRUE),
testSE2<-calcIndividualExpressionsArm(Baseline,PostTreatment,paired=FALSE,min.variance.factor=10^-6)
) 
seMB 
#add NAs and test
testPT<-PostTreatment[1:20,]
testPT<-cbind(rbind(testPT,NaN),NA)
rownames(testPT)[nrow(testPT)]<-"NA"
testB<-Baseline[1:20,]
testB<-cbind(rbind(testB,NaN),NA)
rownames(testB)[nrow(testB)]<-"NA"
#calcIndividualExpressionsC(testB,testPT)) will produce error and stop if NA
```

#Alternate training sets
there is an issue when calling makeComparisons on eset.1 and eset.2 test object, the mclapply is dispatching twice which causes slowness, also I wish to compile R computations for certain functions to speed up before run-time.  This eset was then created from makeCompairson funciton which compares two different labels after splitting the eset by column names label type.  

#Paired end revised demo set , not split by label
```{r}
library(Rcpp)
library(parallel)
library(speedSage)
library(qusage)
eset<-system.file("extdata","eset.RData",package="speedSage")
load(eset)
labels<-c(rep("t0",134),rep("t1",134))
contrast<-"t1-t0"
colnames(eset)<-c(rep("t0",134),rep("t1",134))
fileISG<-system.file("extdata","c2.cgp.v5.1.symbols.gmt",package="speedSage")
ISG.geneSet<-read.gmt(fileISG)
ISG.geneSet<-ISG.geneSet[grepl("DER_IFN_GAMMA_RESPONSE_UP",names(ISG.geneSet))]
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmasCpp.cpp")
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmaArm.cpp")
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmaSingle.cpp")

eset.1<-eset-40.3
eset.2<-eset+100.5
original<-calcIndividualExpressions(eset.1,eset.2,paired=TRUE)
cpp<-calcIndividualExpressionsC(eset.1,eset.2,paired=TRUE)
arm<-calcIndividualExpressionsArm(eset.1,eset.2,paired=TRUE)

e1<-(abs(original[[1]]-arm[[1]]))
e2<-(abs(original[[2]]-arm[[2]]))
e3<-(abs(original[[3]]-arm[[3]]))
e4<-(abs(original[[4]]-arm[[4]]))
qplot(as.vector(e1))
qplot(as.vector(e2))
qplot(as.vector(e3))
qplot(as.vector(e4))


microbenchmark(
 original<-calcIndividualExpressions(eset.1,eset.2,paired=TRUE),
 cpp<-calcIndividualExpressionsC(eset.1,eset.2,paired=TRUE),
 arm<-calcIndividualExpressionsArm(eset.1,eset.2,paired=TRUE))


#showing profiles
library(profr)
library(ggplot2)

yy<-profr(calcIndividualExpressions(eset.1,eset.2,paired=TRUE))
ggplot(yy) + labs(title="Qusage PE Default")
tt<-profr(calcIndividualExpressionsArm(eset.1,eset.2,paired=TRUE))
ggplot(tt)+ labs(title="Qusage PE in Arm")

```





# Non-paired end the eset.1, eset.2 split by label
  This simulates how makeComparison will compare a split eset with label split
```{r}
library(microbenchmark)
library(profr)
library(ggplot2)
library(Rcpp)
eset.1<-system.file("extdata","eset.1.RData",package="speedSage")
eset.2<-system.file("extdata","eset.2.RData",package="speedSage")
load(eset.1)
load(eset.2)
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmasCpp.cpp")
sourceCpp(file="/home/anthonycolombo/Documents/qusage/qusage_repos/qusage_speed/R/sigmaArm.cpp")
original<-calcIndividualExpressions(eset.1,eset.2,paired=FALSE)
cpp<-calcIndividualExpressionsC(eset.1,eset.2,paired=FALSE)
arm<-calcIndividualExpressionsArm(eset.1,eset.2,paired=FALSE)
e1<-(abs(original[[1]]-arm[[1]]))
e2<-(abs(original[[2]]-arm[[2]]))
e3<-(abs(original[[3]]-arm[[3]]))
e4<-(abs(original[[4]]-arm[[4]]))
qplot(as.vector(e1))
qplot(as.vector(e2))
qplot(as.vector(e3))
qplot(as.vector(e4))



microbenchmark(
 original<-calcIndividualExpressions(eset.1,eset.2,paired=FALSE),
 cpp<-calcIndividualExpressionsC(eset.1,eset.2,paired=FALSE),
 arm<-calcIndividualExpressionsArm(eset.1,eset.2,paired=FALSE))


x<-profr(calcIndividualExpressions(eset.1,eset.2,paired=FALSE))
y<-profr(calcIndividualExpressionsArm(eset.1,eset.2,paired=FALSE))
ggplot(x) + labs(title="Qusage SE Default Test 2")
ggplot(y) + labs(title="Qusage SE Armadillo Test 2")


```



