---
title: "Qusage: Speeding in Parallel"
author:  "Timothy J. Triche, Jr, Anthony R. Colombo, Harold Pimentel"
output:
    pdf_document:
      toc: true
      number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

#Make Comparison funciton into parallel
this shows the parallel computation for makeComparison.R in qusage for small test data showing that the changes produce identical results that are faster.


```{r}

```{r}
library(speedSage)
library(qusage)
eset<-system.file("extdata","eset.RData",package="speedSage")
load(eset)
labels<-c(rep("t0",134),rep("t1",134))
contrast<-"t1-t0"
colnames(eset)<-c(rep("t0",134),rep("t1",134))
fileISG<-system.file("extdata","c2.cgp.v5.1.symbols.gmt",package="speedSage")
ISG.geneSet<-read.gmt(fileISG)
ISG.geneSet<-ISG.geneSet[grepl("DER_IFN_GAMMA_RESPONSE_UP",names(ISG.geneSet))]
Baseline<-eset
PostTreatment<-eset+20.4
pairVector<-NULL
var.equal<-FALSE   #case 1 : false variance (no limma)
bayesEstimation<-TRUE
min.variance.factor<-10^-8
paired<-FALSE





